### 2023-03-15 paper.tex
work on methods

@@

Next, we used the single-cell multi-omics data set \citep{kremer_scnmt} to evaluate our DMR detection approach.
We first selected two cell populations from the healthy murine ventricular-subventricular zone: neural stem cells (NSCs, 130 cells) and oligodendrocytes (58 cells).
Then we used the method depicted in Fig.~\ref{fig:dmrscan} to identify DMRs between these two cell types (Fig.~\ref{fig:dmr}).
To estimate the false discovery rate and obtain adjusted p-values, we repeated the DMR detection approach on a permutation of the same data set, generated by shuffling cell labels.
DMRs detected in this manner showed more modest methylation differences, and as a result many DMRs detected in the original data were assigned low adjusted p-values (Fig.~\ref{fig:dmr}A).
Gene ontology (GO) enrichment with GREAT \citep{mclean2010great} revealed that DMRs lowly methylated in oligodendrocytes are located near genes involved in myelination, the main function of oligodendrocytes.
Similarly, DMRs specifically demethylated in NSCs occur near genes involved in stem cell population maintenance.
This demonstrates that our DMR detection approach is able to identify biologically meaningful DMRs, even in scBS data sets of modest cell number.

Fig.~\ref{fig:dmr}C-D depicts an exemplary DMR, located at the gene body of myelin-basic protein (Mbp), the major component of myelin that is essential for myelination of neuronal axons \citep{mbp}.
Our results suggest that oligodendrocyte-specific gene expression of Mbp is supported by low methylation at the detected DMR.

@@

\begin{figure}[h]
    \begin{center}
        \includegraphics[width=0.9\columnwidth]{figures/SFig_benchmark.png}
    \end{center}
    \caption{\small \textbf{Benchmark of our methods on single-cell methylomes of neural subtypes of the murine frontal cortex}.\\
    Our methods were benchmarked on single-cell methylomes from \citet{luo2017single} as described in Fig.~\ref{fig:score}.
    }
    \label{fig:score_luo}
\end{figure}

\begin{figure}[h]
    \begin{center}
        \includegraphics[width=0.9\columnwidth]{figures/SFig_benchmark.png}
    \end{center}
    \caption{\small \textbf{Benchmark of our methods on single-cell methylomes of neural subtypes of the murine frontal cortex}.\\
    Our methods were benchmarked on single-cell methylomes from \citet{luo2017single} as described in Fig.~\ref{fig:score}.
    }
    \label{fig:score_luo}
\end{figure}

@@

\subsection{Re-analysis of published scBS datasets}
To re-analyze scBS data sets by \citet{kremer_scnmt} and \citet{luo2017single}, the genome was first annotated with genomic intervals that were either obtained by tiling the genome into non-overlapping 100~kb tiles, or by acquiring VMRs with \texttt{scbs scan} using the current default options (bandwidth $=2000$, stepsize $=10$, variance threshold $= 2\%$).
For all regions (VMRs and 100~kb tiles), the mean methylation fractions and the shrunken means of the residuals were obtained with \texttt{scbs matrix}.

Regions containing less than four measured CpGs, or that were measured in less than 10\% of cells, were discarded.
To select $n$ intervals for further analysis, the top $n$ intervals were picked based on the number of cells that the region was measured in.

@@

>>>>

Lastly, we assessed whether our proposed workflow requires fine-tuning of parameters (Fig.~\ref{fig:sweep}).
To this end, we re-analyzed both data sets, as well as sub-samples of the data with different VMR detection parameters (the width of the sliding window in bp, \texttt{--bandwidth}, default 2000 and the variance threshold above which windows are merged to VMRs, \texttt{--var-threshold}, default 0.02).
This parameter sweep showed that our workflow is robust to various parameter selections and depends mostly on the overall cell number.
In the more challenging data set (Fig.~\ref{fig:sweep}A), cell types were less cleanly separated when very large bandwidths, or very strict variance thresholds were selected.
However, in our experience small bandwidths or lenient thresholds resulted in a much higher number of VMRs and thus longer computing times.
Our default parameter combination provided good results and fast compute times in both data sets.

@@

Next, we used the single-cell multi-omics data set \citep{kremer_scnmt} to evaluate our DMR detection approach.
We first selected two cell populations from the healthy murine ventricular-subventricular zone: neural stem cells (NSCs, 130 cells) and oligodendrocytes (58 cells).
Then we used the method depicted in Fig.~\ref{fig:dmrscan} to identify DMRs between these two cell types (Fig.~\ref{fig:dmr}).
To estimate the false discovery rate and obtain adjusted p-values, we repeated the DMR detection approach on a permutation of the same data set, generated by shuffling cell labels.
DMRs detected in this manner showed more modest methylation differences, and as a result many DMRs detected in the original data were assigned low adjusted p-values (Fig.~\ref{fig:dmr}A).

Gene ontology (GO) enrichment with GREAT \citep{mclean2010great} revealed that DMRs lowly methylated in oligodendrocytes are located near genes involved in myelination, the main function of oligodendrocytes.
Similarly, DMRs specifically demethylated in NSCs occur near genes involved in stem cell population maintenance.
This demonstrates that our DMR detection approach is able to identify biologically meaningful DMRs, even in scBS data sets of modest cell number.
Fig.~\ref{fig:dmr}C-D depicts an exemplary DMR, located at the gene body of myelin-basic protein (Mbp), the major component of myelin that is essential for myelination of neuronal axons \citep{mbp}.
Our results suggest that oligodendrocyte-specific gene expression of Mbp is supported by low methylation at the detected DMR.

@@

@@

\subsection{Re-analysis of published scBS datasets}
To re-analyze scBS data sets by \citet{kremer_scnmt} and \citet{luo2017single}, the genome was first annotated with genomic intervals that were either obtained by tiling the genome into non-overlapping 100~kb tiles, or by acquiring VMRs with \texttt{scbs scan} using the current default options (bandwidth $=2000$, stepsize $=10$, variance threshold $= 0.02$).
VMRs with sequencing coverage in less than 
% WIP
For all intervals (VMRs and 100~kb tiles), the mean methylation fractions and the shrunken means of the residuals were obtained with \texttt{scbs matrix}.

Regions observed in less than four measured CpGs, or that were measured in less than 10\% of cells, were discarded.
To select $n$ intervals for further analysis, the top $n$ intervals were picked based on the number of cells that the region was measured in.

@@

\begin{figure*}[h]
    \begin{center}
        \includegraphics[width=0.8\columnwidth]{figures/SFig_benchmark.png}
    \end{center}
    \caption{\small \textbf{Benchmark of our methods on single-cell methylomes of neural subtypes of the murine frontal cortex}.\\
    Our methods were benchmarked on single-cell methylomes from \citet{luo2017single} as described in Fig.~\ref{fig:score}.
    }
    \label{fig:score_luo}
\end{figure*}

\begin{figure*}[h]
    \begin{center}
        \includegraphics[width=.6\textwidth]{figures/SFig_parameter_sweep.png}
    \end{center}
    \caption{\small \textbf{Effect of VMR detection parameters on separation of cell types}.\\
    \textbf{(A-B)} Mean neighbor scores obtained after analyzing single-cell methylomes from \citet{kremer_scnmt} (\textbf{A}) or \citet{luo2017single} (\textbf{B}) with our proposed methods.
    VMRs were detected with \texttt{scbs scan} using various sliding window bandwidths or variance thresholds, and on sub-samples of the full data sets.
    }
    \label{fig:sweep}
\end{figure*}
